---
title: "enrichTF-Introduction"
author: "Zheng Wei, Shining Ma, Duren Zhana"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{An Introduction to enrichTF}
  %\VignetteEncoding{UTF-8}
  
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE)
```

# Quick Start


## Download and Installation


The package enrichTF is part of Bioconductor project starting from Bioc 3.9 built on R 3.5. To install the latest version of enrichTF, please check your current Bioconductor version and R version first. The latest version of R is recommended, and then you can download and install enrichTF and all its dependencies as follows:

```{r installpkg,eval=FALSE,message=FALSE,warning=FALSE}
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install("enrichTF")
```

## Loading
Similar with other R packages, please load enrichTF each time before using the package.

```{r loading00,eval=TRUE,message=FALSE}
library(enrichTF)
```

```{r echo=FALSE}
dir.create('refdir')
```

## Use with default configuration

It is quite convenient to run the default pipeline.

Users only need to provide a region list in BED format which contains 3 columns (chr, start, end). It could be the peak calling result of sequencing data like ATAC-seq etc.

All required data and software will be installed automatically.

```{r eval=FALSE}
# Provide your region list in BED format with 3 columns.
foregroundBedPath <- system.file(package = "enrichTF", "extdata","testregion.bed")
# Call the whole pipeline
PECA_TF_enrich(inputForegroundBed = foregroundBedPath, genome = "testgenome") # change"testgenome" to one of "hg19", "hg38", "mm9", 'mm10' ! "testgenome" is only a test example.
```



# Introduction
As transcription factors (TFs) play a crucial role in regulating the transcription process through binding on the genome alone or in a combinatorial manner, TF enrichment analysis is an efficient and important procedure to locate the candidate functional TFs from a set of experimentally defined regulatory regions. While it is commonly accepted that structurally related TFs may have similar binding preference to sequences (i.e. motifs) and one TF may have multiple motifs, TF enrichment analysis is much more challenging than motif enrichment analysis. Here we present a R package for TF enrichment analysis which combine motif enrichment with the PECA model.

# How TF Enrichment Works

## Basic Concept for Relations

In this work, we define four kinds of relations which will be introduced in detail. And we will show how to test on enriched TFs rather than enriched motifs.

### Region-gene Relation Table

When users input a region list in BED format, the relations between regions and one gene can be obtained from two ways:

1.	The regions which are overlapped with the functional region of this gene. This type of relation can be obtained from gene reference data.

2.	The regions are enhancers or other distal regulatory elements of this gene. This type of relation can be derived from 3D genome sequencing experiments such as Hi-C and HiChIP.

As we have already integrated and organized the potential relations between regions and genes from both gene reference and experiment profiles, users can use this source to obtain region-gene relations for their input region list quite easily.

Moreover, users can also customize potential relations from their own work.


### Gene-TF Relation Matrix

The correlation scores between genes and TFs come from our previous approach named PECA. These scores are scaled in the range [-1,1]. Users can also customize this relation matrix based on their own experiment data.


### TF-motif Relation Table

As we know, one TF may relate to multiple motifs. Here, we manually annotated more than 700 motifs with their corresponding TFs.

### Motif-region Relation Table


We scan all annotated motifs on each input region to check if there are some motifs located in this region. Then, the motif-region relation table can be generated.

## Enrichment Test


Based on the foreground region list from users, this package can randomly sample the same number of regions from genome as background.

### Enrichment t-test


Foreground and background region lists can both be connected with genes in Region-gene Relation Table, so both of them will get a gene list. We name them as the foreground gene list and background gene list, respectively.

For each TF, we will perform t-test and obtain p-values for TF-Gene scores of foreground vs those of background. These scores come from Gene-TF Relation Matrix.



### Enrichment Fisher's exact test without threshold


Each TF may connect with many Motifs. For each Motifs this package will do the following fisher exact test to test if this motif correlated enriched in foreground regions with connected genes.

|                                          | fall in region with TF's  motif  | not fall in region with TF's  motif  |
|------------------------------------------|----------------------------|--------------------------------|
| foreground regions (with connected gene) |                            |                                |
| background regions (with connected gene) |                            |                                |

Finally, the package will select the motifs with most significant p-values to represent this TF.

Then, we also can obtain the p-values for all of the TFs


### Enrichment fisher test with threshold

Based on previous section result (the motif selected to represent the TF) and similar to previous section, the package will do the same test on the similar data. The only diffent point is different threshold (for -1 to 1 with interval 0.1) in TF-Gene relation matrix is used to filtered the connected gene. 

In this way, we can select best threshold with the most significant p value. 

Then, we also can get another p value for all of the TF

### FDR

At last, false discover rate correction is apply on last section's result


# Customized Workflow

## Prepare inputs
The function of genBackground is to generate random background regions based on the input regions and to set the length of input sequence regions. We try to select background sequence regions that match the GC-content distribution of the input sequence regions. The default length and number of background regions are 1000 and 10,000, respectively. The length of sequence regions used for motif finding is important. Here, we set 1000 as default value, which means for each region, we get sequences from -500 to +500 relative from center.
For example,
```{r loading01}
setGenome("testgenome")
foregroundBedPath <- system.file(package = "enrichTF", "extdata","testregion.bed")
gen <- genBackground(inputForegroundBed = foregroundBedPath)
```
## Motifscan
The function of MotifsInRegions is to scan for motif occurrences using the prepared PWMs and obtain the promising candidate motifs in these regions.
For example,
```{r loading02}
conTG <- enrichRegionConnectTargetGene(gen)
```
## Connect regions with target genes

The function of regionConnectTargetGene is to connect foreground and background regions to their target genes, which are predicted from PECA model.

For example,
```{r loading03}
findMotif <- enrichFindMotifsInRegions(gen,motifRc="integrate")
```
## TF enrichment test
The function of TFsEnrichInRegions is to test whether each TF is enriched on input regions.
For example,
```{r eval=FALSE}
result <- enrichTFsEnrichInRegions(gen,NULL,NULL)
```


## Build a pipeline

The user can build a pipeline easily based on pipeFrame (pipeline framework)

```{r}
library(magrittr)
setGenome("testgenome")
foregroundBedPath <- system.file(package = "enrichTF", "extdata","testregion.bed")
result <- genBackground(inputForegroundBed = foregroundBedPath) %>%
enrichRegionConnectTargetGene%>%
enrichFindMotifsInRegions(motifRc="integrate") %>%
enrichTFsEnrichInRegions(NULL,NULL)
```

# Session Information

```{r}
sessionInfo()
```
